COVID-19 Vaccine Acceptance and Hesitancy in Low and Middle Income Countries

### 2021-06-22 paper.tex
no google drive
paper
update

  pdftitle={COVID-19 Vaccine Acceptance and Hesitancy in Low and Middle Income Countries, and Implications for Messaging},

@@

@@

@@

@@

\title{COVID-19 Vaccine Acceptance and Hesitancy in Low and Middle Income Countries, and Implications for Messaging}
    \usepackage{authblk}
    \renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
    \renewcommand\Affilfont{\fontsize{8.5}{10.8}\itshape}
                                        \author[1]{Julio S. Solís Arce, BA\footnote{First author}}
                                                            \affil{WZB Berlin Social Science Center}
                                                                                \author[2]{Shana S. Warren, PhD\(^*\)}
                                                            \affil{Innovations for Poverty Action (IPA)}
                                                                                \author[3]{Niccoló F. Meriggi, MSc\(^*\)}
                                                            \affil{International Growth Centre (IGC)}
                                                                                \author[1]{Alexandra Scacco, PhD\(^*\)}
                                                                            \author[1]{Nina McMurry, PhD\(^*\)}
                                                                            \author[4]{Maarten Voors, PhD\(^*\)}
                                                            \affil{Wageningen University \& Research}
                                                                                \author[1,5,24]{Georgiy Syunyaev, MPhil\(^*\)}
                                                            \affil{International Center for the Study of Institutions and Development (HSE University, Moscow, Russia)}
                                                                                \author[6]{Amyn Abdul Malik, PhD\(^*\)}
                                                            \affil{Yale Institute for Global Health}
                                                                                \author[3]{Samya Aboutajdine, MA}
                                                                            \author[7]{Alex Armand, PhD}
                                                            \affil{Nova School of Business and Economics}
                                                                                \author[8]{Saher Asad, PhD}
                                                            \affil{Lahore University of Management Sciences}
                                                                                \author[9]{Britta Augsburg, PhD}
                                                            \affil{The Institute for Fiscal Studies}
                                                                                \author[10]{Antonella Bancalari, PhD}
                                                            \affil{University of St.~Andrews \& The Institute for Fiscal Studies}
                                                                                \author[11]{Martina Björkman Nyqvist, PhD}
                                                            \affil{Stockholm School of Economics and Misum}
                                                                                \author[5,12]{Ekaterina Borisova, PhD}
                                                            \affil{Economics Department of Ghent University}
                                                                                \author[1]{Constantin Manuel Bosancianu, PhD}
                                                                            \author[8,13]{Ali Cheema, PhD}
                                                            \affil{Institute of Development and Economic Alternatives}
                                                                                \author[2]{Elliott Collins, PhD}
                                                                            \author[13]{Ahsan Zia Farooqi, MA}
                                                                            \author[7]{Mattia Fracchia, MA}
                                                                            \author[14]{Andrea Guariso, PhD}
                                                            \affil{Trinity College Dublin}
                                                                                \author[8]{Ali Hasanain, PhD}
                                                                            \author[2]{Anthony Kamwesigye, BSc}
                                                                            \author[15]{Sarah Kreps, PhD}
                                                            \affil{Cornell University}
                                                                                \author[4]{Madison Levine, MSc}
                                                                            \author[16]{Rebecca Littman, PhD}
                                                            \affil{University of Illinois Chicago}
                                                                                \author[17]{Melina Platas, PhD}
                                                            \affil{NYU Abu Dhabi}
                                                                                \author[18]{Vasudha Ramakrishna, MSc}
                                                            \affil{Yale Research Initiative on Innovation and Scale (Y-RISE)}
                                                                                \author[19]{Jacob N. Shapiro, PhD}
                                                            \affil{Princeton University}
                                                                                \author[20]{Jakob Svensson, PhD}
                                                            \affil{Institute for International Economic Studies (IIES), Stockholm University}
                                                                                \author[18]{Corey Vernot, BS}
                                                                            \author[7]{Pedro C. Vicente, PhD}
                                                                            \author[21]{Laurin B Weissinger, DPhil}
                                                            \affil{Tufts University}
                                                                                \author[15]{Baobao Zhang, PhD}
                                                                            \author[2,22]{Dean Karlan, PhD\footnote{Last author}}
                                                            \affil{Kellogg School of Management at Northwestern University}
                                                                                \author[3,23]{Michael Callen, PhD\(^\dagger\)}
                                                            \affil{London School of Economics}
                                                                                \author[3]{Matthieu Teachout, PhD\(^\dagger\)}
                                                                            \author[1,24]{Macartan Humphreys, PhD\(^\dagger\)}
                                                            \affil{Columbia University}
                                                                                \author[6]{Saad B. Omer, PhD\footnote{Corresponding author. 1 Church Street, New Haven, CT 06510. Phone 203 432 3656. Email \href{mailto:saad.omer@yale.edu}{saad.omer@yale.edu}}}
                                                                            \author[25]{Ahmed Mushfiq Mobarak, PhD\footnote{Corresponding author. 165 Whitney Avenue, New Haven, CT 06520-8200. Phone 203 432 5787. Email \href{mailto:ahmed.mobarak@yale.edu}{ahmed.mobarak@yale.edu}}}
                                                            \affil{Yale University}
                                            \date{}

@@

@@

We analyze COVID-19 vaccine acceptance across 15 survey samples covering ten low- and middle-income countries (LMICs) in Asia, Africa, and South America, and two higher income countries (Russia and the United States). Standardized survey responses were collected from `45,928individuals. We find generally high willingness to take a COVID-19 vaccine in LMIC samples (80\% on average), considerably higher than in the United States (65\%) and Russia (30\%). Vaccine acceptance was primarily explained by an interest in personal protection against the disease, while concern about side effects was the most commonly expressed reason for reluctance. Health workers were the most trusted sources of information about COVID-19 vaccines. Our findings suggest that prioritizing efficient and equitable vaccine distribution to LMICs should yield high returns in promoting global immunization. Interventions to promote vaccination in these contexts should focus on translating intentions into uptake and involve trusted health workers.

@@

A safe and effective vaccine against COVID-19 is a critical tool to control the pandemic. As of March 2, 2021, there were 76 vaccines in clinical development and 16 had advanced to Stage 3 clinical trials.\textsuperscript{\protect\hyperlink{ref-who1}{1}} Following clinical trials, several vaccines have been approved in multiple countries and are being distributed across the globe. At present, global vaccine distribution remains highly unequal, with much of the current supply directed toward high-income countries.\textsuperscript{\protect\hyperlink{ref-eiu}{2}}

While effective and equitable distribution of the vaccines is a key policy priority, ensuring the population's acceptance is equally important. Acceptance of vaccines and trust in the institutions that administer them are likely key determinants of the success of any vaccination campaign.\textsuperscript{\protect\hyperlink{ref-defigueiredo2020lancet}{3}} Several studies investigate high-income country residents' willingness to take a potential COVID-19 vaccine.\textsuperscript{\protect\hyperlink{ref-boyon2020ipsos}{4},\protect\hyperlink{ref-Malik2020}{5}} Little is known about acceptance rates in low- and middle-income countries (LMICs), however, where the majority of the world's population resides.

Acceptance of childhood vaccination for common diseases ---such as measles (MCV), Bacille Calmette-Guérin (BCG) and diphtheria, tetanus, and pertussis (DTP)---is generally high in LMICs, providing reasons for optimism about future uptake of COVID-19 vaccines. Table \ref{tab:otherv} summarizes general vaccine acceptance and uptake of common childhood vaccines for the countries included in our study. Still, existing studies on COVID-19 vaccine acceptance document large variation, both across and within countries, including in settings where overall vaccination rates are high.\textsuperscript{\protect\hyperlink{ref-boyon2020ipsos}{4}} These studies highlight concerns about COVID-19 vaccine safety, and particularly concerns about the speed of vaccine development, as reasons for hesitancy in higher income settings. Similar concerns may apply in LMICs.

Additional reasons for hesitancy may feature more prominently in LMIC settings. Reported COVID-19 mortality rates have consistently been lower in LMICs relative to higher income countries.\textsuperscript{\protect\hyperlink{ref-mukherjee}{6}} If individuals in LMICs feel the risk of disease is less serious, they may be less inclined to accept perceived risks associated with taking a recently-developed vaccine. Previous studies have also highlighted factors such as concerns about healthcare quality,\textsuperscript{\protect\hyperlink{ref-christensen2020building}{7}} negative historical experiences,\textsuperscript{\protect\hyperlink{ref-Lowes2018}{8}} weak support from traditional leaders,\textsuperscript{\protect\hyperlink{ref-Jegede2007}{9}} and mistrust in government\textsuperscript{\protect\hyperlink{ref-BLAIR201789}{10}} as barriers to healthcare utilization in LMICs.

Understanding factors that may lead people in LMICs to reject COVID-19 vaccination is of global concern, since a lag in vaccination in the developing world could facilitate the spread of new variants of the virus to other countries.

In order to effectively promote vaccination against COVID-19 and devise messaging strategies, we need to know if people are willing to take COVID-19 vaccines, the reasons why they are willing or unwilling to do so, and the factors influencing their decision-making. For this purpose, we developed and deployed a common set of questions across 15 studies in ten LMICs in Africa (Burkina Faso, Mozambique, Nigeria, Rwanda, Sierra Leone, Uganda), Asia (India, Nepal, Pakistan), and Latin America (Colombia). We compare these findings in LMICs to those from two higher income countries (Russia and the United States).

@@

Our main results are shown in Figure \ref{fig:mainfigure1} and reproduced as Table \ref{tab:maintabledis} in \protect\hyperlink{appendixd}{Appendix A}. The first column provides overall acceptance rates in each study, while the remaining columns show acceptance rates disaggregated by respondent characteristics. The ``All LMICs'' row reports averages for LMIC countries only (and so excludes Russia and the US).

We document meaningful variation in vaccine acceptance across and within LMICs, but generally high levels of acceptance in LMICs overall. The average acceptance across studies is 80.3\% (95\% CI 74.9--85.6), with a median of 78, a range of 30.1 percent points and an interquartile range of 9.7. Our estimate of \(\tau^2\) is 0.007 which implies a standard deviation over country averages of 0.084.

The acceptance rate in every LMIC sample is higher than in USA (64.6\%, 61.8--67.3) and Russia (30.4\%, 29.1--31.7).

We find limited evidence of variation across demographic subgroups in LMICs, as shown in Table \ref{tab:dmeans}. Women were generally less willing to accept the vaccine (average difference about 4.3 points, significant at \(p<.01\)). Younger respondents (defined as aged \textless55 given younger-skewing populations in LMICs) were marginally more willing to take the vaccine, but this difference was not statistically significant. Less educated people were on average more willing to take the vaccine in LMICs, but this difference was not significant.

To better understand the reasoning behind vaccine acceptance, we asked those who were willing to take the vaccine why they would take it. We summarize in Table \ref{tab:yes}, with more details in Table \ref{tab:yesall}.

The most common reason given for vaccine acceptance was personal protection against COVID-19 infection. The average across LMICs was 91\% (86--96). In every individual study, it ranked as the first reason. In distant second place, LMIC respondents reported willingness to take the COVID-19 vaccine in order to protect their families. The average across LMICs was 36\% (28--43). In comparison to self-protection, protecting the community did not feature prominently among the stated reasons.

Self-protection also ranked as the most commonly expressed reason for taking the vaccine in Russia (76\%, 74--78) and the USA (94\%, 92--95).

This evidence contrasts with appeals to altruistic behavior and prosocial motivations in order to promote vaccine acceptance.\textsuperscript{\protect\hyperlink{ref-chou2020considering}{11}} The risks and benefits to personal well-being feature much more prominently in people's stated reasons for vaccine acceptance.

Figure \ref{fig:fig2paper} summarizes the reasons given among respondents who said they were not willing to take a Covid vaccine.

The most common reason expressed for reluctance to take the vaccine in LMIC studies was concern about side effects. For studies Uganda 1 (85.1\% 80.7--89.6), Sierra Leone 2 (57.9\%, 50.1--65.7), Sierra Leone 1 (53.5\% 47.1--59.9) and Uganda 2 (47.3\% 42.2--52.5), more than half of those respondents unwilling to take the vaccine mentioned this reason. Respondents in Russia (36.8\%, 35.2--38.4) and the USA (79.3\%, 74.6--84), reported high levels of this same concern.

While serious adverse events that are life-threatening or require hospitalization are very rare, with only .6\% of respondents reporting at least one side effect in the Pfizer vaccine trial,\textsuperscript{\protect\hyperlink{ref-cdcadverse}{12}} one potential explanation for the outsized concern about side effects could be the lack of widespread information about features of the vaccine at the time of data collection. Media coverage of the few cases of serious adverse events and spread of fake news may contribute as well.\textsuperscript{\protect\hyperlink{ref-stein2017golden}{13}}

Concerns about side effects could also be due to a concern about mild side effects from experiences with other vaccines. In the case of available COVID-19 vaccines, we now know that mild side effects are common but transient. These include fatigue, muscle pain, joint pain and headache, which were severe in fewer than 10\% of people in the clinical trials of tens of thousands. Severe fever occurred in fewer than 2\% of them.\textsuperscript{\protect\hyperlink{ref-wadman2020public}{14}}

Allergic reactions from the COVID-19 vaccine seem to be extremely rare. Data from trials of the Pfizer-BioNTech vaccines shows that anaphylaxis after reported administration occurs at a rate of 11.1 cases per million vaccine doses administered.\textsuperscript{\protect\hyperlink{ref-cdcallergies}{15}} Our data reflects this -- no more than 6\% of respondents expressed concern about allergies in any of our LMIC studies.

Other concerns that make many respondents unwilling to take the vaccine could be countered by accurately presenting the scientific data to the public. Studies Uganda 2 (31\%, 25.9--36.2), Mozambique (29.7\%, 18.6--40.8) and Pakistan 1 (26\%, 18--34) showed relatively high levels of skepticism about vaccine effectiveness. This is also true for respondents in Russia (29.6\%, 28.1--31.1) and the USA (46.8\%, 41--52.6). Recent clinical trials reveal very high rates of vaccine efficacy,\textsuperscript{\protect\hyperlink{ref-baden2021efficacy}{16},\protect\hyperlink{ref-polack2020safety}{17}} and clearly communicating these results to the public is a high priority. In contrast, conspiracy theories were rarely mentioned by respondents in any of our study samples, in spite of widespread popular discourse about anti-vaxxer movements and theories in higher-income countries.\textsuperscript{\protect\hyperlink{ref-loomba_measuring_2021}{18}}

Finally, respondents in some studies downplayed the seriousness of this disease, and listed this as a reason not to be vaccinated. Studies USA (39.3\% 33.5--45), Pakistan 1 (29.4\%, 20.9--37.9) and Nepal (20.4\% 6.7--34.1) reported high rates of respondents citing lack of concern as a reason they were hesitant.

The analysis above identifies the nature of the information gaps that any vaccine messaging should target, while in Figure \ref{fig:fig3paper} we try to identify the actors who are best placed to deliver those messages. We asked respondents about their most trusted source of information when deciding whether to take the vaccine, because these sources are vital to disease control strategies during public health emergencies.\textsuperscript{\protect\hyperlink{ref-siegrist2014role}{19}} Results from Figure \ref{fig:fig3paper} are reproduced as Table \ref{tab:trust} in \protect\hyperlink{appendixd}{Appendix A}.

We find striking consistency across countries. In all but one study, respondents identified the health system as the most trustworthy source to help them decide whether or not to take the COVID-19 vaccine (with the exception of Rwanda, where the government in general was identified as the most trusted source, with the health system a close second). Family and friends were the next most important reference points in most samples. Across samples, women were 3 percentage points more likely to rely on family and friends than male respondents, though this difference was not significant at conventional levels (Figure \ref{fig:genderhist} in \protect\hyperlink{appendixd}{Appendix D}). By contrast, endorsements by religious leaders or celebrity figures were not seen as important sources of influence in any sample other than Nepal.

>>>>

  pdftitle={COVID-19 Vaccine Acceptance and Hesitancy in Low and Middle Income Countries},

@@

\setmainfont{Arial}
\renewcommand{\familydefault}{\sfdefault}

@@

\usepackage{hyperref}

@@

\usepackage{float} \floatplacement{figure}{H}
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

@@

\title{COVID-19 Vaccine Acceptance and Hesitancy in Low and Middle Income Countries}
\author{}
\date{\vspace{-2.5em}}

@@

﻿

Julio S. Solis Arce,\footnote{WZB Berlin Social Science Center, \(^2\)Innovations for Poverty Action (IPA), \(^3\)International Growth Centre (IGC), \(^4\)Wageningen University \& Research, \(^5\)International Center for the Study of Institutions and Development (HSE University, Moscow, Russia), \(^6\)Columbia University, \(^7\)Yale Institute for Global Health, \(^8\)Busara Center for Behavioral Economics, \(^{10}\)Department of Sociology, University of Lagos, \(^{11}\)Busara Nigeria, \(^{12}\)Agricultural and Rural Development Secretariat, Federal Capital Territory Administration (Abuja, Nigeria), \(^{13}\)Nova School of Business and Economics, \(^{14}\)The Institute for Fiscal Studies, \(^{15}\)Lahore University of Management Sciences, \(^{16}\)Innovations for Poverty Action (IPA) Uganda, \(^{17}\)Morsel Research \& Development, \(^{18}\)University of St.~Andrews, \(^{19}\)Redes Peru, \(^{20}\)Stockholm School of Economics and Misum, \(^{21}\)Ghent University, Department of Economics, \(^{22}\)Innovations for Poverty Action (IPA) Colombia, \(^{23}\)Institute of Development and Economic Alternatives, \(^{24}\)Innovations for Poverty Action (IPA) Sierra Leone,\(^{25}\)NOVAFRICA, \(^{26}\)Trinity College Dublin, \(^{27}\)Institute of Public Administration and Management, University of Sierra Leone, \(^{28}\)Centre for the Study of Labour and Mobility (CESLAM), \(^{29}\)Cornell University, \(^{30}\)University of Illinois Chicago, \(^{31}\)Innovations for Poverty Action (IPA) Rwanda, \(^{32}\)Associação NOVAFRICA para o Desenvolvimento Empresarial e Económico de Moçambique, \(^{33}\)Innovations for Poverty Action (IPA) Burkina Faso, \(^{34}\)NYU Abu Dhabi, \(^{35}\)Centre for Economic Research in Pakistan (CERP), \(^{36}\)Yale Research Initiative on Innovation and Scale (Y-RISE), \(^{37}\)Princeton University, \(^{38}\)Institute for International Economic Studies (IIES), Stockholm University, \(^{39}\)Tufts University, \(^{40}\)University of Michigan, \(^{41}\)Kellogg School of Management at Northwestern University, \(^{42}\)London School of Economics, \(^{43}\)Yale University} Shana S. Warren,\(^2\) Niccolò F. Meriggi,\(^3\) Alexandra Scacco,\(^1\) Nina McMurry,\(^1\) Maarten Voors,\(^4\) Georgiy Syunyaev,\(^{1,5,6}\), Amyn Abdul Malik,\(^7\) Samya Aboutajdine,\(^3\) Opeyemi Adeojo,\(^{8, 10}\) Deborah Anigo,\(^{11, 12}\) Alex Armand,\(^{13, 14}\) Saher Asad,\(^{15}\) Martin Atyera,\(^{16}\) Britta Augsburg,\(^{14}\) Manisha Awasthi,\(^{17}\) Gloria Eden Ayesiga,\(^{16}\) Antonella Bancalari,\(^{14, 18, 19}\) Martina Björkman Nyqvist,\(^{20}\) Ekaterina Borisova,\(^{5, 21}\) Constantin Manuel Bosancianu,\(^1\) Magarita Rosa Cabra García,\(^22\) Ali Cheema,\(^{15, 23}\) Elliott Collins,\(^{2}\) Filippo Cuccaro,\(^{24}\) Ahsan Zia Farooqi,\(^{23}\) Tatheer Fatima,\(^{17}\) Mattia Fracchia,\(^{13, 25}\) Mery Len Galindo Soria,\(^{22}\) Andrea Guariso,\(^{26}\) Ali Hasanain,\(^{15}\) Sofía Jaramillo Arciniegas,\(^{22}\) Sellu Kallon,\(^{4, 27}\) Anthony Kamwesigye,\(^{16}\) Arjun Kharel,\(^{28}\) Sarah Kreps,\(^{29}\) Madison Levine,\(^{4}\) Rebecca Littman,\(^{30}\) Shan-e-Mohammad Malik,\(^{23}\) Gisele Manirabaruta,\(^{31}\) Jean Léodomir Habarimana Mfura,\(^{31}\) Fatoma Momoh,\(^{24}\) Alberto Mucauque,\(^{32}\) Imamo Mussa,\(^{32}\) Jean Aime Nsabimana,\(^{31}\) Isaac Obara,\(^{8}\) María Juliana Otálora,\(^{22}\) Béchir Wendemi Ouédraogo,\(^{33}\) Touba Bakary Pare,\(^{33}\) Melina R. Platas,\(^{34}\) Laura Polanco,\(^{22}\) Javaeria Ashraf Qureshi,\(^{30}\) Mariam Raheem,\(^{35}\) Vasudha Ramakrishna,\(^{36}\) Ismail Rendrá,\(^{32}\) Taimur Shah,\(^{35}\) Sarene Eyla Shaked,\(^{16}\) Jacob N. Shapiro, \(^{37}\) Jakob Svensson,\(^{38}\) Ahsan Tariq,\(^{23}\) Achille Mignondo Tchibozo,\(^{33}\) Hamid Ali Tiwana,\(^{23}\) Bhartendu Trivedi,\(^{17}\) Corey Vernot, \(^{36}\) Pedro C. Vicente,\(^{13, 25}\) Laurin B Weissinger,\(^{39}\) Basit Zafar,\(^{35, 40}\) Baobao Zhang,\(^{29}\) Dean Karlan,\(^{2, 41}\) Michael Callen,\(^{42}\) Matthieu Teachout,\(^{3}\) Macartan Humphreys,\(^{1, 6}\) Ahmed Mushfiq Mobarak,\(^{43}\) Saad B. Omer\(^{7}\).

@@

We analyze COVID-19 vaccine acceptance across 15 survey samples covering ten low- and middle-income countries (LMICs) in Asia, Africa, and South America, Russia (an upper-middle-income country), and the United States, including 44,260 individuals. We find considerably higher willingness to take a COVID-19 vaccine in our LMIC samples (mean 80.3\%; median 78\%; range of 30.1\%) compared to the United States (mean 64.6\%) and Russia (mean 30.4\%). Vaccine acceptance is primarily explained by an interest in personal protection against COVID-19, while concern about side effects is the most common reason for hesitancy. Health workers are the most trusted sources of guidance about COVID-19 vaccines. Evidence from this sample of LMICs suggests that prioritizing vaccine distribution to the Global South should yield high returns in advancing global immunization coverage. Vaccination campaigns should focus on translating the high levels of stated acceptance into actual uptake. Messages highlighting vaccine efficacy and safety, delivered by healthcare workers, may be most effective for addressing remaining hesitancy.

@@

A safe and effective vaccine is a critical tool to control the COVID-19 pandemic. As of May 27, 2021, 23 vaccines had advanced to Stage 3 clinical trials\textsuperscript{\protect\hyperlink{ref-who1}{1}} and more than a dozen had been approved in multiple countries\textsuperscript{\protect\hyperlink{ref-track}{2}}. The BioNTech/Pfizer (Comirnaty) vaccine, for example, has been approved in more than 80 countries, while the AstraZeneca vaccine has the most country authorizations at 98.\textsuperscript{\protect\hyperlink{ref-track}{2}} At present, however, global vaccine distribution remains highly unequal, with much of the current supply directed toward high-income countries.\textsuperscript{\protect\hyperlink{ref-wouters2021challenges}{3}}

While effective and equitable distribution of COVID-19 vaccines is a key policy priority, ensuring acceptance is just as important. Trust in vaccines and the institutions that administer them are key determinants of the success of any vaccination campaign.\textsuperscript{\protect\hyperlink{ref-defigueiredo2020lancet}{4}} Several studies have investigated willingness to take a potential COVID-19 vaccine in high-income countries,\textsuperscript{\protect\hyperlink{ref-boyon2020ipsos}{5}--\protect\hyperlink{ref-fisher2020attitudes}{10}} and some studies have included middle-income countries.\textsuperscript{\protect\hyperlink{ref-wouters2021challenges}{3},\protect\hyperlink{ref-lazarus2020nature}{11}} Less is known, however, about vaccine acceptance in low-income countries where large-scale vaccination has yet to begin. Understanding the drivers of COVID-19 vaccine acceptance is of global concern, since a lag in vaccination in any country may result in the emergence and spread of new variants that can overcome immunity conferred by vaccines and prior disease.\textsuperscript{\protect\hyperlink{ref-ong2021lack}{12},\protect\hyperlink{ref-lancettask}{13}}

Our study complements the emerging global picture of COVID-19 vaccine acceptance by focusing primarily on lower income countries. We construct a sample of low- and middle-income countries (LMICs) with wide geographic coverage across Africa, Asia and Latin America. We move beyond documenting vaccine acceptance rates to collect and analyze data on the reasons for acceptance and hesitancy, which is critical for informing the design of effective vaccine distribution and messaging. A summary of the main findings, limitations and implications of the study is shown in Table \ref{tab:summary}.

Acceptance of childhood vaccination for common diseases---such as measles (MCV), Bacille Calmette-Guérin (BCG) and diphtheria, tetanus, and pertussis (DTP)---is generally high in LMICs, providing grounds for optimism about the prospects for COVID-19 vaccine uptake. Table \ref{tab:otherv} summarizes general vaccine acceptance and coverage rates of childhood vaccines in 2018, prior to the current pandemic, for the countries included in our study. Agreement on the importance of childhood vaccinations is markedly higher in the LMICs we study compared to Russia and the United States. Still, existing studies on COVID-19 vaccine acceptance document substantial variation, both across and within countries, including in settings with high acceptance of other vaccinations.\textsuperscript{\protect\hyperlink{ref-wouters2021challenges}{3},\protect\hyperlink{ref-defigueiredo2020lancet}{4},\protect\hyperlink{ref-lazarus2020nature}{11}}

Existing literature cites concern about COVID-19 vaccine safety, including the rapid pace of vaccine development, as a primary reason for hesitancy in higher-income settings.\textsuperscript{\protect\hyperlink{ref-wouters2021challenges}{3},\protect\hyperlink{ref-boyon2020ipsos}{5}} Other reasons may feature more prominently in LMICs. For instance, reported COVID-19 mortality rates have been consistently lower in most LMICs relative to higher income countries.\textsuperscript{\protect\hyperlink{ref-biccard2021patient}{14}--\protect\hyperlink{ref-maeda2021puzzle}{16}} If individuals feel the risk of disease is less severe, they may be less willing to accept any perceived risks of vaccination.\textsuperscript{\protect\hyperlink{ref-brewer2007meta}{17}} Previous studies of healthcare utilization in LMICs have also highlighted factors such as negative perceptions of healthcare quality,\textsuperscript{\protect\hyperlink{ref-christensen2020building}{18}} negative historical experiences involving foreign actors,\textsuperscript{\protect\hyperlink{ref-Lowes2018}{19},\protect\hyperlink{ref-martinez2021vaccines}{20}} weak support from traditional leaders,\textsuperscript{\protect\hyperlink{ref-Jegede2007}{21}} and mistrust in government\textsuperscript{\protect\hyperlink{ref-BLAIR201789}{22}} as barriers to uptake, which could apply to COVID-19 vaccination as well.

To promote vaccination against COVID-19, we need to know whether people are willing to take COVID-19 vaccines, the reasons why they are willing or unwilling to do so, and the most trusted sources of information in their decision-making. Our study investigates these questions using a common set of survey items deployed across 13 studies in Africa, South Asia, and Latin America: seven in low-income countries (Burkina Faso, Mozambique, Rwanda, Sierra Leone, Uganda), five in lower-middle-income countries (India, Nepal, Nigeria, Pakistan), and one in an upper-middle-income country (Colombia). We compare these findings to those from two countries at the forefront of vaccine research and development, Russia and the United States.

To select studies to include in our sample, we conducted an internal search within Innovations for Poverty Action (IPA), the International Growth Center (IGC), and the Berlin Social Science Center (WZB) for projects with plans to collect survey data in the second half of 2020. Study PIs agreed to include a set of common questions about COVID-19 vaccine attitudes. This strategy was guided by the need to collect information quickly and cost-effectively using a survey modality (phone) that was both safe, given pandemic conditions, and appropriate for contexts with limited internet coverage. The final set of samples included in our study therefore reflects populations that fall under the current research priorities at IGC, IPA and WZB and, in the case of IPA and IGC, donors that prioritize working in the Global South.

@@

Our main results are shown in Figure \ref{fig:mainfigure1} and reproduced as \hyperref[tab:maintabledis]{Supplementary Table 1}. The first column provides overall acceptance rates in each study, while the remaining columns disaggregate acceptance by respondent characteristics. The ``All LMICs'' row reports averages for the LMIC samples included in our study and excludes Russia and the USA. The ``LMICs National Samples'' row reports averages for just the LMIC samples with national-level geographic coverage.

The average acceptance rate across the full set of LMIC studies is 80.3\% (95\% CI 74.9--85.6\%), with a median of 78\%, a range of 30.1\% and an interquartile range of 9.7\%. Our estimate of the between-study standard deviation, \(\tau\), using a random effects meta-analysis model is 0.084 which represents only 10.5\% of our estimate of the average acceptance across LMIC studies.

The acceptance rate in every LMIC sample is higher than in the USA (64.6\%, CI 61.8--67.3\%) and Russia (30.4\%, CI 29.1--31.7\%). Reported acceptance is lowest in Burkina Faso (66.5\%, CI 63.5--69.5\%) and Pakistan 2 (66.5\%, CI 64.1--68.9\%). Pakistan's relatively low acceptance rate may be linked to negative historical experiences with foreign-led vaccination campaigns.\textsuperscript{\protect\hyperlink{ref-martinez2021vaccines}{20},\protect\hyperlink{ref-ali2019polio}{23},\protect\hyperlink{ref-robbins2012cia}{24}} This hesitancy may be particularly problematic given the magnitude of the second wave in neighboring India and acceleration of cases across South Asia that threaten to overwhelm health infrastructure. The relatively low acceptance rate in Burkina Faso may reflect general vaccine hesitancy. As shown in Table \ref{tab:otherv}, fewer people believe that vaccines in general are safe in Burkina Faso than in any other country included in our study except Russia.

We find limited evidence of variation across demographic subgroups in our aggregate analysis of LMIC samples, as shown in \hyperref[tab:dmeans]{Supplementary Table 7}. Women are generally less willing to accept the vaccine than men (average difference about 4.2 points, significant at \(p<.01\)). Respondents under age 25 and less educated respondents are marginally more willing to take the vaccine compared to older and more educated respondents, respectively, but these differences are not statistically significant.

Table \hyperref[tab:countrydiff]{Supplementary Table 9} provides results disaggregated by demographic subgroups for individual studies. The average gender differences in the aggregate LMIC analysis are driven by the Burkina Faso, Mozambique, Pakistan 1, Rwanda and Sierra Leone 1 samples. However, these gender differences in acceptance are less than 10 percentage points in each of these samples, in contrast to larger gender gaps in acceptance we observe in the USA (17\%) and Russia (16\%).

Less-educated respondents expressed significantly higher acceptance in the Burkina Faso, Rwanda, Sierra Leone 1, and Uganda 2 samples, which represent the majority of studies from sub-Saharan Africa. Notably, we observe the opposite pattern in the India, Pakistan 1, and Pakistan 2 samples. In all three studies, acceptance is greater among more educated respondents, although this difference is not statistically significant in the India sample. Education is also a positive and significant predictor of acceptance in the USA.

We find mixed evidence across studies with respect to the relationship between age and COVID-19 vaccine acceptance. In India and Nigeria, respondents younger than 25 years old are significantly less willing to take the vaccine relative to adults 25-54 years old, while in Mozambique, Pakistan 1, and Rwanda, those under 25 are significantly more willing. In Mozambique and Rwanda, respondents under 25 are also significantly more accepting compared to those 55 and over; however, the difference between these age groups is not statistically significant in other LMIC samples. In the USA and Russia, older respondents are consistently more accepting than younger respondents.

To better understand the reasoning behind vaccine acceptance, we asked those who were willing to take the vaccine why they would take it. We summarize these results in \hyperref[tab:yes]{Supplementary Table 2}, with additional details in \hyperref[tab:yesall]{Supplementary Table 3}. The reason most commonly given for vaccine acceptance across samples is personal protection against COVID-19 infection. The average across the LMIC samples is 91\% (CI 86--96\%) with a median of 92.5\% and a range of 22\%. In every individual study, including the USA (94\%, CI 92--95\%) and Russia (76\%, CI 74--78\%), this ranks as the most cited reason. In distant second place in the aggregate LMIC analysis is family protection, with an average of 36\% (CI 28--43\%), a median of 34.5\% and a range of 39\%. In comparison to protecting oneself and one's family, protecting one's community does not feature prominently among stated reasons for acceptance. These reasons do not vary substantially by age group, as shown in \hyperref[tab:yes1]{Supplementary Table 4}.

Figure \ref{fig:fig2paper} summarizes the reasons given by respondents who said they were not willing to take a COVID-19 vaccine. Concern about side effects is the most frequently expressed reason for reluctance in our LMIC samples. This concern is particularly evident among samples from sub-Saharan Africa. In studies Uganda 1 (85.1\%, CI 80.7--89.6\%), Sierra Leone 2 (57.9\%, CI 50.1--65.7\%), Sierra Leone 1 (53.5\%, CI 47.1--59.9\%) and Uganda 2 (47.3\%, CI 42.2--52.5\%), more than half of respondents unwilling to take the vaccine cited worries about side effects. Respondents in Russia (36.8\%, CI 35.2--38.4\%) and even more in the USA (79.3\%, CI 74.6--84\%), frequently report this same concern.

Study samples Uganda 2 (31\%, CI 25.9--36.2\%), Mozambique (29.7\%, CI 18.6--40.8\%) and Pakistan 1 (26\%, CI 18--34\%) show relatively high levels of skepticism about vaccine effectiveness among hesitant respondents. This is also true in Russia (29.6\%, CI 28.1--31.1\%) and the USA (46.8\%, CI 41--52.6\%). In addition, some hesitant respondents cite lack of concern about COVID-19 infection as a reason not to be vaccinated. This answer is particularly common in the USA (39.3\%, CI 33.5--45\%), Pakistan 1 (29.4\%, CI 20.9--37.9\%) and Nepal (20.4\%, CI 6.7--34.1\%) studies.

In Figure \ref{fig:fig3paper} we report respondents' most trusted source of guidance when deciding whether to take a COVID-19 vaccine. Results from Figure \ref{fig:fig3paper} are reproduced in \hyperref[tab:trust]{Supplementary Table 6}, while \hyperref[tab:q4]{Supplementary Table 13} presents a complete description of response recoding from individual studies.

We find striking consistency across studies. In all samples except Rwanda, including those from Russia and the USA, respondents identify the health system as the most trustworthy source to help them decide whether to take the COVID-19 vaccine. The average across LMIC samples is 48.1\% (CI 31.6--64.5\%), with a median of 44.1\% and a range of 66.3\%. Respondents in Sierra Leone 2 (89.3\%, CI 87.2--91.5\%), Nigeria (58\%, CI 55.7--60.2\%) and Burkina Faso (51.6\%, CI 48.5--54.8\%) cited health workers most often. Sierra Leone has the highest trust in health workers and the Ministry of Health, potentially reflecting investments in public health following the 2014-2015 Ebola epidemic.\textsuperscript{\protect\hyperlink{ref-deserrano}{25}}

In Colombia (36.6\%, CI 33.5--39.7\%), Nepal (35.6\%, CI 32.9--38.3\%), Russia (28.1\%, CI 26.8--29.3\%) and Burkina Faso (18.4\%, CI 16--20.9\%), the next most-cited sources are family and friends. Across the pooled samples, women are 3 percentage points more likely to rely on family and friends than male respondents, though this difference is not statistically significant (\hyperref[fig:genderhist]{Supplementary Figure 1}).
